---
title: "Data Frame Exercises"
author: "Srikanth Srinivas"
date: "`r Sys.Date()`"
output: html_document
---

# Data Frame Creation and Challenges

In this document, we will create a data frame with five columns and perform various operations on the data.

## Step 1: Install and Load Required Packages

To begin, we need to install and load the necessary packages. (I have installed them already, so they're commented out.)

```{r, eval=FALSE}
# install.packages("randomNames")
# install.packages("wakefield")
```
```{r}
library(randomNames) # For generating random names
library(wakefield)   # For generating random ages and sexes
```
##Step 2: Create the Data Frame
Now, we will generate the data and create the data frame with 5 columns: First_Name, Last_Name, Age, Sex, and Test_Score.
```{r}
# Generate random names
first_names <- randomNames(40, which.names = "first")
last_names <- randomNames(40, which.names = "last")

# Generate ages between 14 and 56 using sample
ages <- sample(14:56, 40, replace = TRUE)

# Generate sexes
sexes <- sex_inclusive(40)

# Generate test scores between 0 and 100
test_scores <- sample(0:100, 40, replace = TRUE)

# Combine into a data frame
student_data <- data.frame(
  First_Name = first_names,
  Last_Name = last_names,
  Age = ages,
  Sex = sexes,
  Test_Score = test_scores
)

# View the first few rows of the data frame
head(student_data)
```
#Step 4: Challenge 2 - Summarizing Data
Now, we will calculate the mean and standard deviation of the test scores for each sex.
```{r}
# Calculate mean and standard deviation of scores for each sex
summary_stats <- aggregate(Test_Score ~ Sex, data = student_data, 
                            FUN = function(x) c(mean = mean(x), sd = sd(x)))

# Convert to data frame for better readability
summary_stats_df <- do.call(data.frame, summary_stats)

# View the summary statistics
summary_stats_df
```
#Step 5: Challenge 3 - Adding Logical Vectors
Next, we will add a new column called "Pass," which is TRUE if the score is greater than or equal to 70, and FALSE otherwise.
```{r}
# Add "Pass" column
student_data$Pass <- student_data$Test_Score >= 70

# Calculate the percentage of students who passed
pass_percentage <- mean(student_data$Pass) * 100

# Display the percentage of students who passed
pass_percentage
```
#Step 6: Challenge 4 - Multi Filtering
Finally, we will filter the data frame to select only rows where Age is between 25 and 40, Sex is female, and Score is more than 80.
```{r}
# Multi-filter the data frame
filtered_data <- student_data[student_data$Age > 25 & student_data$Age < 40 & 
                              student_data$Sex == "Female" & 
                              student_data$Test_Score > 80, ]

# View the filtered data
filtered_data
```

